---
title: "2022_new_fix_east"
author: "Xuelei_Zhang"
date: '2023-03-07'
output: html_document
---

New file with fit_ssm and fit_mpm used different time steps hence more locations 

also convert longitude from 0-180 to 0-360

3/12/2023 ISSUE: some locations in 2021 data have the same move persistence estimates value. also large sd errors. (package version: aniMotum 1.1-0.4)

Maybe go back to older version animotum (1.1) OR foieGras (0.7_6)



```{r}


library(plyr)
library(aniMotum)
library(ggplot2)
library(lubridate)
library(argosfilter)
library(tidyr)
library(dplyr)
library(viridis)
library(sf)
library(tictoc)
library(rnaturalearth)
library(plotly)


world_map <- map_data("world")

sc <- scale_colour_gradientn(colours = viridis(100), limits=c(0,1))


```


```{r}

# load raw data 

#load("G:/auckland/nz_right_whale/maps/analysis/data processing/new/raw_argos_df.Rdata")
load("D:/auckland/nzsrw/maps/maps/analysis/data processing/new/raw_argos_df.Rdata")
 
# skip to SDA filter 

raw_argos_df <- ddply(raw_argos_df, ~id, function(d){
      d$argosfilter <- sdafilter(lat = d$lat, 
                               lon = d$lon, 
                               lc = d$lc, 
                               dtime = d$date, vmax = 25)
        return(d)
})

head(raw_argos_df)

# visualize filter 

with(raw_argos_df, plot(lon,lat,col="lightgrey",type="p",xlab="Longitude",ylab="Latitude", pch=19, cex=0.8, xlim=c(55,170), ylim=c(-65, -35))) #all data
with(raw_argos_df, points(lon[which(argosfilter=="not")],lat[which(argosfilter=="not")],col="blue", pch=19, cex=0.8)) #all data to be retained
with(raw_argos_df, points(lon[which(argosfilter=="removed")],lat[which(argosfilter=="removed")],col="red", pch=19, cex=0.5)) #removed locations

# exclude errorneous locs 
filtered_argos_df <- raw_argos_df %>%
  filter(argosfilter != "removed") %>%
  dplyr::select(-argosfilter)



filtered_argos_df %>% 
  group_by(id) %>% 
  dplyr::summarize(nb_locations = n())

tab_1 <- raw_argos_df %>% 
  group_by(id) %>% 
  dplyr::summarize(nb_locations = n())

tab_2 <- filtered_argos_df %>% 
  group_by(id) %>% 
  dplyr::summarize(nb_locations = n())

tab  <- plyr::join(data.frame(tab_1), data.frame(tab_2), by="id")
colnames(tab) <- c("id", "raw_locs", "filt_locs")
tab


```


```{r}


pre_dup <- nrow(filtered_argos_df) # to get the current number of data points

# create dummy variable
filtered_argos_df$index <- c(1:nrow(filtered_argos_df))

# run each tag in a loop check to check for duplicates
# if there is a time duplicate, select the best quality position or simply the first position
filtered_argos_df <- ddply(filtered_argos_df, ~id, function(d){
  toremove <- c()
  for (i in c(2:nrow(d))) {
    if (d$date[i] == d$date[i-1]) {
      dd <- d[(i-1):i,]
      r <- dd[dd$lc == ave(dd$lc, FUN = min), "index"] # find the lowest quality
      toremove <- c(toremove, r[1]) #select first element of r in case both locations have the same lq
    }
  }
  if (length(toremove) > 0){d <- d[!(d$index %in% toremove), ]}
  return(d)
})
# remove dummy variable
filtered_argos_df$index <- NULL
pre_dup - nrow(filtered_argos_df)





```


```{r}


time_diff_hours_df <- ddply(filtered_argos_df, ~id, function(d){
  d$time_diff_hours <- NA
  for (i in 2:nrow(d)){
    d$time_diff_hours[i] = as.numeric(difftime(d$date[i], d$date[i-1], units = "hours"))}
  return(d)
})

# mean time difference between locations (in hours)

mts <- aggregate(time_diff_hours~id, time_diff_hours_df, mean)

mts #this is the mean time step

mxts <- aggregate(time_diff_hours~id, time_diff_hours_df, max)
mxts# this is the max time step

mnts <- aggregate(time_diff_hours~id, time_diff_hours_df, min)
mnts # this is the minimum time step

mets <- aggregate(time_diff_hours~id, time_diff_hours_df, median)
mets # this is the median time step


ggplot(time_diff_hours_df, aes(time_diff_hours)) + geom_histogram(binwidth = 1, col ="white", na.rm = T) + theme_bw() + xlim(c(0,100)) + xlab("Time difference between successive locations")

ggplot(time_diff_hours_df, aes(time_diff_hours)) + geom_histogram(binwidth = 1, col ="white", na.rm = T) + theme_bw() + xlim(c(0,15)) + xlab("Time difference between successive locations")


```

```{r}

trackseg_argos_df <- ddply(time_diff_hours_df, ~id, function(d){
ind <- which(d$time_diff_hours > 24)
d$mark <- 0
d$mark[ind] <- 1
d$track_seg <- cumsum(d$mark)
  return(d)
})

# Now create a new id based on track segment
trackseg_argos_df$track_id <- paste(trackseg_argos_df$id, "-", trackseg_argos_df$track_seg, sep="")

table(trackseg_argos_df$track_id)

length(unique(trackseg_argos_df$track_id)) #96

# remove short track segs n <10 

min_obs <- 10 ## set the number of minimum obs acceptable
trackseg_argos_df <- trackseg_argos_df %>% group_by(track_id)
trackseg_argos_df_filt <- filter(trackseg_argos_df, n() >= min_obs)

table(trackseg_argos_df_filt$track_id)
length(unique(trackseg_argos_df_filt$track_id)) # 63 

# 42387 locs 

```

```{r}

ssm_df <- trackseg_argos_df_filt[,c(2:5,9)]

ssm_df <- ssm_df%>% filter (lc!="Z")

ssm_tdiff_hours_df <- ddply(ssm_df, ~track_id, function(d){
  d$time_diff_hours <- NA
  for (i in 2:nrow(d)){
    d$time_diff_hours[i] = as.numeric(difftime(d$date[i], d$date[i-1], units = "hours"))}
  return(d)
})

mts <- aggregate(time_diff_hours~ track_id, ssm_tdiff_hours_df, mean)
mts 

mean(mts$time_diff_hours) # 3.4 hrs for overall 

ssm_df <- ssm_df[,c(5,1:4)] #structure the data frame so it matches the required structure
colnames(ssm_df)[1] <- "id"



ssm_df <- data.frame(ssm_df)

#save(ssm_df,file="D:/auckland/nzsrw/maps/maps/analysis/data processing/new/ssm_df_3.8.Rdata")

```


```{r ssm 2022}

# td = 2.67h
# pre fit_ssm = 9022
# after: 3109
ssm_22 <- ssm_df %>% filter(date > "2022-07-01")

ssm_22<- ssm_22  %>% subset (id != "215262-18"&
                            id != "215262-20"&
                            id!="215262-22"&
                            id != "215258-9")
                          
table(ssm_22$id)
length(unique(ssm_22$id))


fit_ssm_6_22 <- fit_ssm(ssm_22,vmax = 25,model="crw",time.step = 6,control = ssm_control(verbose=0))

print(fit_ssm_6_22)

ssm_22_df <- grab(x=fit_ssm_6_22,what = "p")

#aniMotum::map(fit_ssm_6_22,what = "predicted",by.date=F,aes = aes_lst(mp=FALSE,fill=c("dodgerblue", "dodgerblue", NA, "orange", "grey60", NA),conf = F,date_pal = hcl.colors(100, palette = "Viridis", rev = FALSE)))


# move persistence model 

mpm_22 <- fit_mpm(fit_ssm_6_22,model="mpm",control = mpm_control(verbose = 0))

mpm_22

mpm_22_df <- grab(x=mpm_22,what = "f") 

#plot(mpm_22, control = mpm_control(verbose=0),ask=F)



ssm_mpm_22 <- aniMotum::join(fit_ssm_6_22,mpm_22,what.ssm="predicted",as_sf=FALSE)%>%
               as.data.frame()


ggplot(ssm_mpm_22, aes(lon, lat)) +
  geom_point(size=1, aes(col = g)) +
  geom_polygon(data = world_map, aes(x=long, y=lat, group=group), fill="black") +
 coord_equal() + 
  coord_fixed(xlim=c(160,180), ylim=c(-56,-45))+
  theme_bw()+
  theme(panel.grid=element_blank())+
  sc
ggplotly()

mean(ssm_mpm_22$g)
#save(ssm_mpm_22_df,file="G:/auckland/nz_right_whale/maps/analysis/data processing/new/2022/ssm_mpm_22.Rdata")

save(ssm_mpm_22,file="D:/auckland/nzsrw/maps/maps/analysis/data processing/new/2022/ssm_mpm_22.Rdata")

```

average time difference for 215262 = 6.1h
average time difference without 215262 = 2.8h
overall td = 4.19h

```{r ssm 2021}

# 15990 lcs without 215262

ssm_21 <- ssm_df %>% subset(    id == "46633-0"| 
                                   
                                 id == "46635-0"| 
                                 id == "46635-2"| 
                                 id == "46635-3"| 
                          
                                 id == "46950-0"|
                                  
                                 id == "46955-0"|
                                 
                                 id == "212499-0"|
                                   
                                 id == "212500-0"|
                                   
                                 id == "215258-0"|
                                 id == "215258-1"|
                                 id == "215258-3"|
                                 id == "215258-4"|
                                 id == "215258-9"|
                                   
                                 id == "215259-0"|
                                 id == "215259-1"|
                                 id == "215259-2"|
                                 id == "215259-3"|
                                 id == "215259-5"|
                                   
                                 id == "215261-0"|
                                 id == "215261-2"|
                                  
                                 #id == "215262-0"|
                                 #id == "215262-1"|
                                 #id == "215262-2"|
                                 #id == "215262-4"|
                                 #id == "215262-6"|
                                 #id == "215262-7"|
                                 #id == "215262-9"|
                                 #id == "215262-12"|
                                 #id == "215262-13"|
                                 #id == "215262-14"|
                                 #id == "215262-15"|
                                 #id == "215262-16"|
                                 #id == "215262-18"|
                                # id == "215262-20"|
                                 #id =="215262-22"|
                                  
                                 id == "215263-0")

length(unique(ssm_21$id))


fit_ssm_6_21 <- fit_ssm(ssm_21,vmax = 25,model="crw",time.step =6,control = ssm_control(verbose=0))


#plot(fit_ssm_6_21, what = "predicted", type = 2, ask = F,alpha=0.1)


ssm_21_df <- grab(x=fit_ssm_6_21,what = "p") # 5h 5694/6h 4748/ 8h 3574

# change longitude to 0-360 
mpm_21<- ssm_21_df %>%
                select(id,date,lon,lat)%>%
               mutate(lon = lon %% 360) %>%
               fit_mpm(., model = "mpm")
mpm_21

plot(mpm_21,ask=F)

# 46635-2 215258-9 did not converge 
mpm_21_df <- grab(x=mpm_21,what = "f") 

ssm_mpm_21 <- aniMotum::join(fit_ssm_6_21,mpm_21,what="p")%>%
              as.data.frame()

ssm_mpm_21 <- ssm_mpm_21 %>% filter (id != "46635-2"&
                                      id != "215258-9")
           
ggplot(data.frame(ssm_mpm_21), aes(lon, lat)) +
  geom_point(size=1, aes(col = g)) +
  geom_polygon(data = world_map, aes(x=long, y=lat, group=group), fill="black") +
 coord_equal() + 
  coord_fixed(xlim=c(160,180), ylim=c(-56,-45))+
  theme_bw()+
  theme(panel.grid=element_blank())+
  sc
ggplotly()

mean(ssm_mpm_21$g)
#4723 lcs without 215262

```

```{r 2021 215262}

ssm_215262 <- ssm_df %>% subset( id == "215262-0"|
                                 id == "215262-1"|
                                 id == "215262-2"|
                                 id == "215262-4"|
                                 id == "215262-6"|
                                 id == "215262-7"|
                                 id == "215262-9"|
                                 id == "215262-12"|
                                 id == "215262-13"|
                                 id == "215262-14"|
                                 id == "215262-15"|
                                 id == "215262-16"|
                                 id == "215262-18"|
                                 id == "215262-20"|
                                 id =="215262-22")


fit_ssm_215262 <- fit_ssm(ssm_215262,vmax = 25,model="crw",time.step =11,control = ssm_control(verbose=0))

ssm_215262_df <- grab(x=fit_ssm_215262,what = "p") 

#mpm
mpm_215262 <- fit_mpm(fit_ssm_215262,model="mpm",control = mpm_control(verbose = 0))

mpm_215262

mpm_215262_df <- grab(x=mpm_215262,what = "f") 

table(ssm_215262_df$id)

plot(mpm_215262,ask=F)


ssm_mpm_215262 <- aniMotum::join(fit_ssm_215262,mpm_215262,what.ssm="predicted",as_sf=FALSE)                    %>% as.data.frame()


# 439 lcs

ssm_mpm_21 <- rbind (ssm_mpm_21,ssm_mpm_215262)

```


```{r}
# final ssm_mpm_21 5162 lcs
ssm_mpm_21 <- rbind (ssm_mpm_21,ssm_mpm_215262)

save(ssm_mpm_21,file="D:/auckland/nzsrw/maps/maps/analysis/data processing/new/2021/ssm_mpm_21.Rdata")

```


2020

203571 8/6/2020 10:00:00 PM 
203572 8/6/2020 9:00:00 PM 
203573 8/5/2020 1:00:00 AM 
203574 8/16/2020 8:00:00 PM 
203575 8/4/2020 9:00:00 PM 
205015 8/6/2020 10:00:00 PM

```{r 2020}


# 16121 lcs

ssm_20 <- ssm_df %>% subset (id == "203571-0"|
                                  id == "203571-1"|
                                    
                                  id == "203572-0"|
                                  id == "203572-1"|
                                    
                                  id == "203573-0"|
                                  id == "203573-1"|
                                    
                                  id == "203574-0"|
                                    
                                  id == "203575-0"|
                                  id == "203575-2"|
                                  id == "205015-0")
# average time difference = 1.914

table(ssm_20$id)

fit_ssm_5_20<- fit_ssm(ssm_20, vmax=25, model="crw", time.step=5, control = ssm_control(verbose=0))

fit_ssm_5_20

#plot(fit_ssm_6_20,ask=F,type=2,alpha=0.1,what="p")


ssm_20_df <- grab(fit_ssm_5_20,what="p") # 4503

table(ssm_20_df$id)

#ssm_20_df <- ssm_20_df %>% filter (id != "203573-1")

mpm_20<- fit_mpm(fit_ssm_5_20,model="jmpm",control = mpm_control(verbose = 0))

mpm_20

#plot(mpm_20,ask=F)


mpm_20_df <- grab(x=mpm_20,what = "f") 

ssm_mpm_20<- dplyr::full_join(x=ssm_20_df,y=mpm_20_df)

ggplot(data.frame(ssm_mpm_20),aes(lon, lat)) +
  geom_point(size=1, aes(col = g)) +
  geom_polygon(data = world_map, aes(x=long, y=lat, group=group), fill="black") +
 coord_equal() + 
  coord_fixed(xlim=c(160,180), ylim=c(-56,-45))+
  theme_bw()+
  theme(panel.grid=element_blank())+
  sc
ggplotly()

mean(ssm_mpm_20$g)

# fit_mpm jmpm mean g= 0.7 ts=4h 5622 locs 
# mpm did not fit when ts = 4h 

# ts = 5h 4503 locs; NaNs produced "203573-1 didnot converge when used mpm 
# ts = 5h jmpm mean=0.66
# 5h seems to be better lowe CI move persistence lines seem to be smoother 
```


```{r add name for 2020}

ssm_mpm_20$Name = ssm_mpm_20$id

ssm_mpm_20$Name[ssm_mpm_20$Name=="203571-0"] <- "Rima"
ssm_mpm_20$Name[ssm_mpm_20$Name=="203571-1"] <- "Rima"


ssm_mpm_20$Name[ssm_mpm_20$Name=="203572-0"] <- "Toru"
ssm_mpm_20$Name[ssm_mpm_20$Name=="203572-1"] <- "Toru"


ssm_mpm_20$Name[ssm_mpm_20$Name=="203573-0"] <- "Rua"
ssm_mpm_20$Name[ssm_mpm_20$Name=="203573-1"] <- "Rua"

ssm_mpm_20$Name[ssm_mpm_20$Name=="203574-0"] <- "Wiremu-Bill"

ssm_mpm_20$Name[ssm_mpm_20$Name=="203575-0"] <- "Tahi"
ssm_mpm_20$Name[ssm_mpm_20$Name=="203575-2"] <- "Tahi"

ssm_mpm_20$Name[ssm_mpm_20$Name=="205015-0"] <- "Whaa"

save(ssm_mpm_20,file="D:/auckland/nzsrw/maps/maps/analysis/data processing/new/2020/ssm_mpm_20.Rdata")

```